BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2017-03-30 Paper2/BOAST.tex
Modified title.
fix and improve SPECFEM3D modifications

In this last subsection, we present an alternative use-case of the BOAST
framework, focused on portability. As part of the Mont-Blanc project,
we had to port a scientific application so that it could be used to
benchmark the Mont-Blanc HPC cluster. It was composed of 40 complex
and hand-tuned CUDA kernels, and we had to port them to OpenCL. This
scenario also highlights the factorization and maintainability
improvements of the BOAST framework, as we obtained a single BOAST
implementation of the kernels (instead of CUDA plus OpenCL) and this
version of the code is still used and updated in the upstream project.

@@

Because of the complexity of wave propagation (and of the application
architecture), it is hard to impossible to define unit tests. Hence,
the application is validated by the accuracy of the results it
produces, that is, the accuracy of seismograms of the simulated
earthquakes, in comparison with the actual ones. Non-regression
testing (after new developments) is also based on these seismograms,
with measurements of the relative error between two identical
simulations.

As part of the Mont-Blanc project, we had to port SPECFEM3D to OpenCL, so
that it could be used to benchmark Mont-Blanc HPC platforms.

@@

%%%
\paragraph{Porting run-time to OpenCL} The second part of the port
consisted in the translation of the CPU-side of the application, from
CUDA API to OpenCL API. Most of the functions of the interfaces are
very similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of OpenCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations, and 270 releases.

@@

%%%
\paragraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
merged in SPECFEM3D's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and 24 Intel
Xeon processors, we measured identical performance between the original
CUDA version and our BOAST-CUDA version. This result was expected, as
the BOAST-generated version is identical to the original, apart from
naming differences.

@@

By refactoring GPU kernels' code in BOAST EDSL, the size of the kernels' source code
shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, mainly
because of the removal of code duplication and manually unrolled
loops). This is beneficial for SPECFEM3D as it improves the readability
and maintainability of its source code. This gain was confirmed by the
SPECFEM3D community as the subsequent developments have been carried out
on the BOAST version of the kernels, and not on the CUDA or OpenCL generated code.

>>>>

In this last subsection, we present an alternative use-case of the BOAST
framework, focused on portability. As part of the Mont-Blanc project,
we ported a scientific application so that it could be use to
benchmark the Mont-Blanc HPC cluster. It was composed of 40 complex
and hand-tuned CUDA kernels, that we ported to OpenCL framework. This
scenario also highlights the factorization and maintainability
improvements of the BOAST framework, as we obtained a single BOAST
implementation of the kernels (instead of two, CUDA and OpenCL). This
version of the code is still used and updated in the upstream project.

@@

Because of the complexity of the wave propagation kernels, it was
impossible for us to understand the kernel source kernel. Hence, we
treated them as black-boxes, and aimed at obtaining identical
performances. Furthermore, the kernels' complexity prevents the
specification of unit tests, thus the application is validated by the
accuracy of the results it produces (the seismograms of the simulated
earthquakes), against the actual ones. Non-regression testing (after
new developments) is also based on these seismograms, with
measurements of the relative error between two identical simulations.

@@

%%%
\paragraph{Porting run-time to OpenCL} The second part of the port
consisted in the translation of the CPU-side of the application, from
CUDA API to OpenCL'. Most of the functions of the interfaces are very
similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of OpenCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations, and 270 releases.

@@

%%%
\paragraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
merged in SPECFEM3D's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and 24 Intel
Xeon processors, we measured identical performance between the original
CUDA version and our BOAST-CUDA version. This result was expected, as
the BOAST-generated version is identical to the original, except from
naming differences.

@@

Thanks to the porting of the SPECFEM3D GPU kernels to BOAST EDSL, the size of the
kernels' source code shrank by a factor of 1.8 (from 7500 to less than
4000 LOC, mainly because of the removal of code duplication and
manually unrolled loops). This is beneficial for SPECFEM3D as it
improves the readability and maintainability of its source code. This
gain was confirmed by the SPECFEM3D community as the subsequent
developments have been carried out on the BOAST version of the
kernels, and not on the CUDA or OpenCL generated code.
